### 2021-02-15 Chapter-3/Chapter-3.tex
update

\section{Math Notations}

\section{Scalability}

\section{Evaluation}

>>>>

\section{Math Notations}
e now discuss how to model observations ($O$), which are the direct information
that an adversary can get during the attack.

During the execution, a program ($\beta$) have many temporary values ($t_i \in
T$). Once $\beta$ (program), $k$ (secret), and $m$ (message, public) are
determined, $t_i$ is also fixed. Therefore, $ t_i = f_i(\beta, k, m)$, where $f_
i$ is a function that maps between $t_i$ and ($\beta$, $k$, $m$).

In the paper, we consider two code patterns that can be exploited by an attacker,
\emph{secret-dependent control transfers} and \emph{secret-dependent data
accesses}. In other words, an adversary has observations based on control-flows
and data accesses.

\subsection{Secret-dependent Control Transfers}
We think a control-flow is secret-dependent if different input sensitive keys
($K$) can lead to different branch conditions. Therefore,
We define a branch is secret-dependent if:
$$\exists k_{i1}, k_{i2} \in K, \,f_i(\beta, k_{i1}, m) \neq f_i(\beta, k_{i2}, m)$$

An adversary can observe which branch the code executes, if the branch condition
equals to $t_b$. We use the constraint $c_i : f_i(\beta, k, m) = t_b$ to model
the observation ($o$) on secret-dependent control-transfers.

\subsection{Secret-dependent Data Accesses}
Similar to secret-dependent control-flow transfers, a data access operation is
secret-dependent if different input sensitive keys ($K$) can lead to different
memory addresses. We use the model from CacheD~\cite{203878}. The low $L$ bits
of the address are irrelevant in side-channels.

We consider a data access is secret-dependent if:
$$\exists k_{i1}, k_{i2} \in K, \,f_i(\beta, k_{i1}, m) >> L \neq f_i(\beta, k_{i2}, m) >> L$$

If the memory access equals to $t_b$, we can use the constraint $c_i :
f_i(\beta, k, m) >> L = t_b >> L$ to model the observation on secret-dependent
data accesses.

With the above definitions, we can model an attacker's observation by a series of math
formulas. For example, in Figure~\ref{fig:side-channel}, if an attacker observes
the code executes the branch 1, we have $c_5: k_1 + k_2 < 4$ to describe an
attacker's knowledge and $K^{o5} = \{k_1,\, k_2\,|\, (k_1 + k_2) < 4\}$. If an
attacker observes the code executes the branch 2, we have $c_8: k_1 - k_2 > 0$
and $K^{o8} = \{k_1,\, k_2\,|\, (k_1 - k_2) > 0\}$.
\section{Scalability}
\subsection{Trace-oriented Symbolic Execution}
While symbolic execution can capture fine-grained semantics of programs, it
is also notorious for its unbearable performance cost. Previous trace-oriented
symbolic execution based
works~\cite{203878,Chattopadhyay:2017:QIL:3127041.3127044} all have large
performance bottlenecks. As a result, those approaches either only apply to
small-size programs~\cite{Chattopadhyay:2017:QIL:3127041.3127044} or apply some
domain knowledge to simplify the analysis. 
%Those tools interpret each
%instruction and update memory cells and registers with formulas that
%captured the semantics of the execution and search different input values that
%can lead to different execution behaviors using constraint solver. 
We implement
the approach presented in \S\ref{sec:trace-qif} and model the side-channels as
formulas. While the tool can finish analyzing some simple cases like AES, it can
not handle complicated cases like RSA.
We observe that finding side-channels using symbolic execution is different from
traditional general symbolic execution and can be optimized to be as efficient
as other methods with approaches below.

\subsection{Interpret Instructions Symbolically}
Existing binary analysis tools~\cite{shoshitaishvili2016state,
10.1007/978-3-642-22110-1_37} usually translate machine instructions into
intermediate languages (IR) to simplify analysis. 
The reason is that the number of machine instructions is
enormous, and the semantics of each instruction is complex. Intel Developer
Manual~\cite{intelsys} introduces more than 1000 different x86 instructions. 
IR typically has fewer instructions compared to the original machine ISA\@.
However, the IR layer, which predigest the implementation
and reduce the workload of those tools, is not suitable for side-channels 
analysis. Memory-based side-channels are very low issues. So, in general,
IR-based or source code side-channels analyses are not accurate enough.
In many cases, compilers can use conditional moves or bitwise operations to eliminate
branches. Also, as some IRs are not a superset or a subset of ISA, 
it is hard to rule out conditional jumps introduced by IR and add real branches 
eliminated by IR transformations.

Moreover, the IR design introduces significant overhead~\cite{217563}.
Transferring machine instructions into IR is time-consuming. For example,
REIL IR~\cite{dullien2009reil}, adopted in CacheS~\cite{236338}, has multiple
transform processes, from binary to VEX IR, BAP IR, and finally REIL IR\@. 
Also, IR increases the total number of instructions. For example, x86
instruction \textit{test eax, eax} transfers into 18 REIL IR instructions. If we
assume the time of symbolically executing grows as the number of instructions increase, the
design of adopting IR layers can introduce large overhead.

\vspace*{2pt}
\textbf{Our Solution to Challenge C2:}
We adopt the approach from QSYM~\cite{217563} and implement the symbolic execution
directly on the top of x86 instructions. Table~\ref{scala:ir} shows that
eliminating the IR layer can reduce the number of instructions executed during
the analysis.

\begin{table}%[ht]
      \centering\small\footnotesize
      \caption{The number of x86,  % instructions and the number of 
             REIL IR, and VEX IR instructions on the traces of crypto programs.}
      \label{scala:ir}
\vspace*{-1pt}
      \resizebox{\columnwidth}{!}{%

            \begin{tabular}{cccc}
                  \hline
                                    & \begin{tabular}[c]{@{}c@{}}Number of\\ x86 Instructions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Number of\\ VEX IR\end{tabular} & \begin{tabular}[c]{@{}c@{}}Number of\\ REIL IR\end{tabular} \\ \hline
                  AES OpenSSL 0.9.7 & $1,704$                   & $23,938$ (15x)            & $62,045$ (36x)            \\
                  DES OpenSSL 0.9.7 & $2,976$                   & $41,897$ (15x)            & $100,365$ (33x)           \\
                  RSA OpenSSL 0.9.7 & $1.6*10^7$                & $2.4*10^8$ (15x)          & $5.9*10^8$ (37x)          \\
                  RSA mbedTLS 2.5  & $2.2*10^7$                & $3.1*10^8$ (15x)          & $8.6*10^8$  (39x)         \\ \hline
            \end{tabular}
      }
\end{table}

\subsection{Constraint Solving}
As discussed in \S\ref{side-channel:condition}, the problem of identifying
side-channels can be reduced to the question below.

\begin{quote}
      \textit{Can we find two different input variables $k_1, k_2 \in K$ that
            satisfy the formula $f_a(k_1) \neq f_a(k_2)$?}
\end{quote}

Existing approach relies on satisfiability modulo theories (SMT) solvers (e.g,
Z3~\cite{DeMoura:2008:ZES:1792734.1792766}) to find satisfying $k_1$ and $k_2$.
We argue that while it is a universal approach to solving constraints with SMT
solvers, for constraints with the above formats, using custom heuristics and
testing is much more efficient in practice. Constraint solving is a decision
problem expressed in logic formulas. SMT solvers transfer the inputted SMT
formula into the boolean conjunctive normal form (CNF) and feed it into the
internal boolean satisfiability problem (SAT) solver. The translation process,
called ``bit blasting'', is time-consuming. Also, as the SAT problem is a
well-known NP-complete problem, it is also hard to deal when it comes to
realistic uses with huge formulas. Despite the rapid development of SMT solvers
in recent years, constraint solving remains one of the obstacles to achieve the
scalability for real-world cryptosystems.

\vspace*{2pt}
\textbf{Our Solution to Challenge C3:}
Instead of feeding the formula $f_a(k_1) \neq f_a(k_2)$ into a SMT solver, we
just randomly pick up $k_1, k_2 \in K$ and test them if they can satisfy the
formula. Our solution is based on the following intuition. For most combination
of $(k_{1}, k_{2} )$, the formula $f_a(k_1) \neq f_a(k_2)$ holds. As long as
$f_a$ is not a constant function, such $k_1, k_2$ must exist. For example,
suppose each time we only have 5\% chance to find such $k_1, k_2$, then after we
test with different input combination with 100 times, we have $1 -
(1-0.05)^{100} = 99.6\%$ chance find such $k_1, k_2$. Such random algorithms
work well for our problem.
\section{Evaluation}
